---
title: "Behaviours"
author: "Gabriel Melhado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
pacman::p_load(car, dplyr, ggplot2, scales, here, tidyr, tidyverse, lme4,lmerTest,emmeans, performance, broom.mixed, DHARMa, ggpubr, psych, glmmTMB)
```

```{r}
# Read dataset
data <- read.csv("dataset.csv")
```


# Exploratory Data Analysis
```{r}
#Data structure
glimpse(data)
summary(data)
colSums(is.na(data))
```

```{r}
#Descriptive analysis
data %>%
  group_by(treatment) %>%
  reframe(across(where(is.numeric),
                 \(x) c(mean = mean(x, na.rm = TRUE),
                        sd = sd(x, na.rm = TRUE),
                        median = median(x, na.rm = TRUE),
                        min = min(x, na.rm = TRUE),
                        max = max(x, na.rm = TRUE))))
```


```{r}
# Checking data distribution

summary(data$distance_cm)
summary(data$freezing_s)
summary(data$refuge_duration_s)
summary(data$velocity_cm.s)


hist(data$distance_cm)
hist(data$freezing_s)
hist(data$refuge_duration_s)
hist(data$velocity_cm.s)


```


```{r}
#Densities - B
data %>%
  pivot_longer(cols = where(is.numeric),
               names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value, fill = treatment)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~ Variable, scales = "free") +
  theme_bw()
```

```{r}
#Boxplots
data %>%
  pivot_longer(cols = where(is.numeric),
               names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = treatment, y = Value, fill = treatment)) +
  geom_boxplot(outlier.color = "red", alpha = 0.7) +
  facet_wrap(~ Variable, scales = "free_y") +
  theme_bw()

```

## Data Analysis - Behaviours
# Distance_moved: the total distance (in cm) that fish swim during the behavioural trial. Tank_ID was firstly included in the analysis as a random factor, but the data indicated that, for all behavioural responses, tanks do not explain any addtional variation beyond the residual error.

```{r}
# Prepare variables
data$treatment <- factor(data$treatment,
                       levels = c("control", "low", "high"))

data$sex <- factor(data$sex)
data$tank_id <- factor(data$tank_id)
data$body_mass_sc <- scale(data$body_mass)

# Linear model
m_distance <- lm(
  distance_cm ~ treatment + sex + body_mass_sc,
  data = data
)

summary(m_distance)

# Diagnosis
res_m1_distance_moved <- simulateResiduals(m_distance)
plot(res_m1_distance_moved)

testDispersion(res_m1_distance_moved)
testZeroInflation(res_m1_distance_moved)
testOutliers(res_m1_distance_moved)

#Anova of the model
Anova(m_distance, type = "II")

emmeans(m_distance, pairwise ~ treatment, adjust = "tukey")

```
## Plotting
# Behaviour: Distance_moved

```{r}
#---------------------------------------------------------
# 1) Ensure treatment order
#---------------------------------------------------------

data$treatment <- factor(
  data$treatment,
  levels = c("control", "low", "high")
)

# Labels for plotting
data$treatment_plot <- factor(
  data$treatment,
  levels = c("control", "low", "high"),
  labels = c("Control", "Low", "High")
)

#---------------------------------------------------------
# 2) Fit final model
#---------------------------------------------------------

m_distance <- lm(
  distance_cm ~ treatment + sex + body_mass,
  data = data
)

summary(m_distance)

Anova(m_distance, type = "II")

#---------------------------------------------------------
# 3) Estimated marginal means
#---------------------------------------------------------

emm_dist_df <- as.data.frame(
  emmeans(m_distance, ~ treatment)
)

print(emm_dist_df)

#---------------------------------------------------------
# 4) Standardise column names
#---------------------------------------------------------

cn <- names(emm_dist_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

if (is.na(mean_col) | is.na(lower_col) | is.na(upper_col)) {
  stop("Could not identify CI columns.")
}

emm_dist_df <- emm_dist_df %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# 5) Colour palette
#---------------------------------------------------------

pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)

#---------------------------------------------------------
# 6) Plot
#---------------------------------------------------------

p_distance <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = distance_cm,
      fill = treatment_plot
    ),
    trim = FALSE,
    alpha = 0.35,
    colour = NA
  ) +

  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = distance_cm,
      colour = treatment_plot
    ),
    width = 0.18,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_dist_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = treatment_plot
    ),
    width = 0.25,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_dist_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    limits = c(0, 3000),
    breaks = seq(0, 3000, by = 500),
    labels = comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Distance moved (cm)"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    plot.title = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_distance)

#---------------------------------------------------------
# 7) Save
#---------------------------------------------------------

# ggsave(here("figures", "distance_moved.pdf"), p_distance, width = 6, height = 5)

# ggsave(here("figures", "distance_moved.jpg"), p_distance, width = 6, height = 5, dpi = 600)
```

# Time_freezing: time (in s) that fish was in open arena but not moving during the trials.

```{r}
# Linear model

m_freezing <- lm(
  freezing_s ~ treatment + sex + body_mass_sc,
  data = data
)

summary(m_freezing)

#Diagnosis
res_m1_freezing <- simulateResiduals(m_freezing)
plot(res_m1_freezing)

testDispersion(res_m1_freezing)
testZeroInflation(res_m1_freezing)
testOutliers(res_m1_freezing)


#Anova
Anova(m_freezing, type = "II")
emmeans(m_freezing, pairwise ~ treatment, adjust = "tukey")

```

## Plotting
# Behaviour: Time_freezing

```{r}
#---------------------------------------------------------
# Estimated marginal means
#---------------------------------------------------------

emm_freezing_df <- as.data.frame(
  emmeans(m_freezing, ~ treatment)
)

print(emm_freezing_df)

#---------------------------------------------------------
# Standardise column names
#---------------------------------------------------------

cn <- names(emm_freezing_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

if (is.na(mean_col) | is.na(lower_col) | is.na(upper_col)) {
  stop("Could not identify CI columns.")
}

emm_freezing_df <- emm_freezing_df %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes
#---------------------------------------------------------

pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)

#---------------------------------------------------------
# Plot
#---------------------------------------------------------

p_freezing <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = freezing_s,
      fill = treatment_plot
    ),
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = freezing_s,
      colour = treatment_plot
    ),
    width = 0.18,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_freezing_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = treatment_plot
    ),
    width = 0.25,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_freezing_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    limits = c(0, 1200),
    breaks = seq(0, 1200, by = 200),
    labels = scales::comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Freezing duration (s)"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    plot.title = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_freezing)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

# ggsave( filename = here("figures", "freezing_duration.pdf"), plot = p_freezing, width = 6, height = 5)

# ggsave(filename = here("figures", "freezing_duration.jpg"), plot = p_freezing, width = 6, height = 5, dpi = 600)

```

# Refuge_duration: time that fish spend under the shelter during the trials.

# Refuge duration exhibited a zero-inflated continuous distribution, with approximately 13% of fish never entering the refuge (0 s). To appropriately model this response, a Tweedie GLM was fitted because the Tweedie distribution accommodates continuous positive values while allowing an exact mass at zero, making it well suited for ecological time-to-event and duration data with excess zeros. 

```{r}
#Model: family Tweedie
m_refuge <- glmmTMB(
    refuge_duration_s ~ treatment + sex + body_mass_sc,
    family = tweedie(link = "log"),
    data = data
)

summary(m_refuge)

# Diagnosis

simulationOutput <- simulateResiduals(m_refuge)
plot(simulationOutput)
testDispersion(simulationOutput)
testZeroInflation(simulationOutput)
testOutliers(simulationOutput)

#Anova of the model
Anova(m_refuge, type = "II")
emm_refuge <- emmeans(
    m_refuge,
    ~ treatment,
    type = "response"
)

emm_refuge

pairs(emm_refuge, adjust = "tukey")
```

## Plotting
#Behaviour: Shelter use

```{r}
#---------------------------------------------------------
# Estimated marginal means (response scale)
#---------------------------------------------------------

emm_refuge_df <- as.data.frame(
  emmeans(
    m_refuge,
    ~ treatment,
    type = "response"
  )
)

print(emm_refuge_df)

#---------------------------------------------------------
# Standardise column names
#---------------------------------------------------------

cn <- names(emm_refuge_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_refuge_df <- emm_refuge_df %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes
#---------------------------------------------------------

pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)

#---------------------------------------------------------
# Plot
#---------------------------------------------------------

p_refuge <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = refuge_duration_s,
      fill = treatment_plot
    ),
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = refuge_duration_s,
      colour = treatment_plot
    ),
    width = 0.18,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_refuge_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = treatment_plot
    ),
    width = 0.25,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_refuge_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    limits = c(0, 1200),
    breaks = seq(0, 1200, by = 200),
    labels = scales::comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Refuge use duration (s)"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    plot.title = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_refuge)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

# ggsave(filename = here("figures", "refuge_duration.pdf"), plot = p_refuge, width = 6, height = 5)

# ggsave(filename = here("figures", "refuge_duration.jpg"), plot = p_refuge, width = 6, height = 5, dpi = 600)
```

# Velocity

```{r}
# Linear model
m_velo <- lm(
  velocity_cm.s ~ treatment + sex + body_mass_sc,
  data = data
)

summary(m_velo)

# Diagnosis
res_m1_velo <- simulateResiduals(m_velo)
plot(res_m1_velo)

testDispersion(res_m1_velo)
testZeroInflation(res_m1_velo)
testOutliers(res_m1_velo)

#Anova of the model
Anova(m_velo, type = "II")

emmeans(m_velo, pairwise ~ treatment, adjust = "tukey")
```
# Plotting velocity
```{r}
#---------------------------------------------------------
# Estimated marginal means from the main model
#---------------------------------------------------------

emm_velo_df <- as.data.frame(
  emmeans(
    m_velo,
    ~ treatment
  )
)

emm_velo_df <- emm_velo_df %>%
  mutate(
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  )

#---------------------------------------------------------
# Plot
#---------------------------------------------------------

p_velocity <- ggplot() +

  # Violin
  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = velocity_cm.s,
      fill = treatment_plot
    ),
    alpha = 0.18,
    colour = NA,
    trim = FALSE,
    width = 0.85
  ) +

  # Raw observations
  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = velocity_cm.s,
      colour = treatment_plot
    ),
    width = 0.16,
    alpha = 0.70,
    size = 2
  ) +

  # Estimated means
  geom_errorbar(
    data = emm_velo_df,
    aes(
      x = treatment_plot,
      ymin = lower.CL,
      ymax = upper.CL,
      colour = treatment_plot
    ),
    width = 0.20,
    linewidth = 1.2
  ) +

  geom_point(
    data = emm_velo_df,
    aes(
      x = treatment_plot,
      y = emmean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_fill_manual(values = pastel_cols) +
  scale_colour_manual(values = dark_cols) +

  labs(
    x = "Fluoxetine treatment",
    y = expression("Velocity (cm s"^{-1}*")")
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_velocity)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

ggsave(
  here("figures", "velocity.pdf"),
  p_velocity,
  width = 6,
  height = 5
)

ggsave(
  here("figures", "velocity.jpg"),
  p_velocity,
  width = 6,
  height = 5,
  dpi = 600
)
```



## Interaction Behaviours x Sex: A treatment × sex interaction was included to test whether the behavioural response to fluoxetine differed between males and females. This is biologically relevant because sex-specific differences in immune activation and sickness behaviour have been reported in vertebrates and may influence responses to environmental contaminants.

# Behaviour: Distance_moved

```{r}
# Linear model
m_distance_int <- lm(
  distance_cm ~ treatment * sex + body_mass_sc,
  data = data
)

summary(m_distance_int)

# Diagnosis
res_m1_distance_moved_int <- simulateResiduals(m_distance_int)
plot(res_m1_distance_moved_int)

testDispersion(res_m1_distance_moved_int)
testZeroInflation(res_m1_distance_moved_int)
testOutliers(res_m1_distance_moved_int)

#Anova of the model
Anova(m_distance_int, type = "III")

emmeans(m_distance_int, pairwise ~ treatment | sex, adjust = "tukey")

emmeans(m_distance_int, pairwise ~ sex | treatment, adjust = "tukey")



```

# Plotting Distance_moved x Sex

```{r}
#---------------------------------------------------------
# Prepare labels
#---------------------------------------------------------

data <- data %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

#---------------------------------------------------------
# Estimated marginal means for treatment × sex interaction
#---------------------------------------------------------

emm_distance_int_df <- as.data.frame(
  emmeans(
    m_distance_int,
    ~ sex | treatment
  )
)

cn <- names(emm_distance_int_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_distance_int_df <- emm_distance_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes for sex
#---------------------------------------------------------

sex_pastel_cols <- c(
  "Female" = "#FDD0A2",
  "Male"   = "#C6DBEF"
) 

sex_dark_cols <- c(
  "Female" = "#E6550D",
  "Male"   = "#2171B5"
) 

#---------------------------------------------------------
# Plot: sex side by side within treatment
#---------------------------------------------------------

pd <- position_dodge(width = 0.75)

pj <- position_jitterdodge(
  jitter.width = 0.12,
  dodge.width = 0.75
)

p_distance_int_side <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = distance_cm,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = distance_cm,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_distance_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_distance_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    limits = c(0, 4000),
    breaks = seq(0, 4000, by = 500),
    labels = scales::comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Distance moved (cm)",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_distance_int_side)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

ggsave(
  filename = here("figures", "distance_moved_treatment_sex_side_by_side.pdf"),
  plot = p_distance_int_side,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "distance_moved_treatment_sex_side_by_side.jpg"),
  plot = p_distance_int_side,
  width = 7,
  height = 5,
  dpi = 600
)
```

# Time-freezing x sex:

```{r}
# Linear model

m_freezing_int <- lm(
  freezing_s ~ treatment * sex + body_mass_sc,
  data = data
)

summary(m_freezing_int)

#Diagnosis
res_m1_freezing_int <- simulateResiduals(m_freezing_int)
plot(res_m1_freezing_int)

testDispersion(res_m1_freezing_int)
testZeroInflation(res_m1_freezing_int)
testOutliers(res_m1_freezing_int)


#Anova
Anova(m_freezing_int, type = "III")
emmeans(m_freezing_int, pairwise ~ treatment | sex, adjust = "tukey")
emmeans(m_freezing_int, pairwise ~ sex | treatment, adjust = "tukey")

```

#Plottin time_freezing x sex

```{r}
#---------------------------------------------------------
# Prepare labels
#---------------------------------------------------------

data <- data %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

#---------------------------------------------------------
# Estimated marginal means for treatment × sex interaction
#---------------------------------------------------------

emm_freezing_int_df <- as.data.frame(
  emmeans(
    m_freezing_int,
    ~ sex | treatment
  )
)

cn <- names(emm_freezing_int_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_freezing_int_df <- emm_freezing_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes for sex
#---------------------------------------------------------

sex_pastel_cols <- c(
  "Female" = "#FDD0A2",
  "Male"   = "#C6DBEF"
) 

sex_dark_cols <- c(
  "Female" = "#E6550D",
  "Male"   = "#2171B5"
) 

#---------------------------------------------------------
# Plot: sex side by side within treatment
#---------------------------------------------------------

pd <- position_dodge(width = 0.75)
pj <- position_jitterdodge(
  jitter.width = 0.12,
  dodge.width = 0.75
)

p_freezing_int_side <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = freezing_s,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = freezing_s,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_freezing_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_freezing_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    limits = c(0, 1200),
    breaks = seq(0, 1200, by = 200),
    labels = scales::comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Freezing duration (s)",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_freezing_int_side)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

ggsave(
  filename = here("figures", "freezing_treatment_sex_side_by_side.pdf"),
  plot = p_freezing_int_side,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "freezing_treatment_sex_side_by_side.jpg"),
  plot = p_freezing_int_side,
  width = 7,
  height = 5,
  dpi = 600
)

```


#Shelter_use * Sex

```{r}
#Model: family Tweedie
m_refuge_int <- glmmTMB(
    refuge_duration_s ~ treatment * sex + body_mass_sc,
    family = tweedie(link = "log"),
    data = data
)

summary(m_refuge_int)

# Diagnosis

simulationOutput <- simulateResiduals(m_refuge_int)
plot(simulationOutput)
testDispersion(simulationOutput)
testZeroInflation(simulationOutput)
testOutliers(simulationOutput)

#Anova of the model
Anova(m_refuge_int, type = "III")

emm_refuge_int <- emmeans(
    m_refuge_int,
    ~ treatment | sex,
    type = "response"
)

pairs(
    emm_refuge_int,
    adjust = "tukey"
)

emmeans(
    m_refuge_int,
    pairwise ~ sex | treatment,
    type = "response",
    adjust = "tukey"
)

```

#Plotting shleter_use * sex

```{r}
#---------------------------------------------------------
# Prepare labels
#---------------------------------------------------------

data <- data %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

#---------------------------------------------------------
# Estimated marginal means for treatment × sex interaction
# response scale because model uses Tweedie with log link
#---------------------------------------------------------

emm_refuge_int_df <- as.data.frame(
  emmeans(
    m_refuge_int,
    ~ sex | treatment,
    type = "response"
  )
)

cn <- names(emm_refuge_int_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_refuge_int_df <- emm_refuge_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes for sex
#---------------------------------------------------------

sex_pastel_cols <- c(
  "Female" = "#FDD0A2",
  "Male"   = "#C6DBEF"
) 

sex_dark_cols <- c(
  "Female" = "#E6550D",
  "Male"   = "#2171B5"
) 

#---------------------------------------------------------
# Plot: sex side by side within treatment
#---------------------------------------------------------

pd <- position_dodge(width = 0.75)

pj <- position_jitterdodge(
  jitter.width = 0.12,
  dodge.width = 0.75
)

p_refuge_int_side <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = refuge_duration_s,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = refuge_duration_s,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_refuge_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_refuge_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    limits = c(0, 1200),
    breaks = seq(0, 1200, by = 200),
    labels = scales::comma
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Refuge use duration (s)",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_refuge_int_side)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

ggsave(
  filename = here("figures", "refuge_treatment_sex_side_by_side.pdf"),
  plot = p_refuge_int_side,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "refuge_treatment_sex_side_by_side.jpg"),
  plot = p_refuge_int_side,
  width = 7,
  height = 5,
  dpi = 600
)
```


#Velocity * sex
```{r}
# Linear model
m_velo_int <- lm(
  velocity_cm.s ~ treatment * sex + body_mass_sc,
  data = data
)

summary(m_velo_int)

# Diagnosis
res_m1_velo_int <- simulateResiduals(m_velo_int)
plot(res_m1_velo_int)

testDispersion(res_m1_velo_int)
testZeroInflation(res_m1_velo_int)
testOutliers(res_m1_velo_int)

#Anova of the model
Anova(m_velo_int, type = "III")
emmeans(m_velo_int, pairwise ~ treatment | sex, adjust = "tukey")
emmeans(m_velo_int, pairwise ~ sex | treatment, adjust = "tukey")

```

#Plotting velocity * sex
```{r}
#---------------------------------------------------------
# Prepare labels
#---------------------------------------------------------

data <- data %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

#---------------------------------------------------------
# Estimated marginal means for treatment × sex interaction
#---------------------------------------------------------

emm_velo_int_df <- as.data.frame(
  emmeans(
    m_velo_int,
    ~ sex | treatment
  )
)

cn <- names(emm_velo_int_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_velo_int_df <- emm_velo_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Colour palettes for sex
#---------------------------------------------------------

sex_pastel_cols <- c(
  "Female" = "#FDD0A2",
  "Male"   = "#C6DBEF"
) 

sex_dark_cols <- c(
  "Female" = "#E6550D",
  "Male"   = "#2171B5"
) 

#---------------------------------------------------------
# Plot
#---------------------------------------------------------

pd <- position_dodge(width = 0.75)

pj <- position_jitterdodge(
  jitter.width = 0.12,
  dodge.width = 0.75
)

p_velocity_int_side <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = velocity_cm.s,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = velocity_cm.s,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_velo_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_velo_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    limits = c(0, 4),
    breaks = seq(0, 4, by = 0.5)
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Velocity (cm/s)",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_velocity_int_side)

#---------------------------------------------------------
# Save
#---------------------------------------------------------

ggsave(
  filename = here("figures", "velocity_treatment_sex_side_by_side.pdf"),
  plot = p_velocity_int_side,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "velocity_treatment_sex_side_by_side.jpg"),
  plot = p_velocity_int_side,
  width = 7,
  height = 5,
  dpi = 600
)
```


