---
title: "Body condition"
author: "Gabriel Melhado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
pacman::p_load(car, dplyr, ggplot2, scales, here, tidyr, tidyverse, lme4,lmerTest,emmeans, performance, broom.mixed, DHARMa, ggpubr, psych, glmmTMB)
```


```{r}
# Read dataset
data <- read.csv("dataset.csv")
```


## Checking sex ratio per treatment

```{r}
# Sex ratio per treatment
sex_ratio_treatment <- data %>%
  group_by(treatment, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = sex,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(
    total = female + male,
    prop_female = female / total,
    prop_male = male / total,
    sex_ratio = paste0(male, ":", female)
  )

sex_ratio_treatment
```

## Checking sex ratio per fish tank

```{r}
# Sex ratio per fish tank
sex_ratio_tank <- data %>%
  group_by(treatment, tank_id, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = sex,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(
    total = female + male,
    prop_female = female / total,
    prop_male = male / total,
    sex_ratio = paste0(male, ":", female)
  ) %>%
  arrange(treatment, tank_id)

sex_ratio_tank

```


```{r}
# Prepare variables
data <- data %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    sex = factor(
      sex,
      levels = c("female", "male")
    ),
    tank_id = factor(tank_id),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

```


```{r}
# Create long-format dataset for repeated body mass

mass_long <- data %>%
  dplyr::select(
    fish_id,
    tank_id,
    treatment,
    sex,
    body_length,
    body_mass_before,
    body_mass_after
  ) %>%
  pivot_longer(
    cols = c(body_mass_before, body_mass_after),
    names_to = "time",
    values_to = "body_mass"
  ) %>%
  mutate(
    time = factor(
      time,
      levels = c("body_mass_before", "body_mass_after"),
      labels = c("Before LPS", "After LPS")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

summary(mass_long$body_mass)

```

## Tank variance = 0 and model is singular. Therefore, tank_id was removed from the random-effects structure.
### Testing full model

```{r}
# Mixed model: body mass before vs after LPS

m_mass_time <- lmer(
  body_mass ~ time * treatment * sex + (1 | fish_id),
  data = mass_long,
  REML = FALSE
)

summary(m_mass_time)
isSingular(m_mass_time)

par(mfrow = c(2,2))
plot(m_mass_time)
qqnorm(residuals(m_mass_time))
qqline(residuals(m_mass_time))
hist(residuals(m_mass_time))

Anova(m_mass_time, type = "III")

emm_mass_time <- emmeans(
  m_mass_time,
  ~ time | treatment
)

emm_mass_time

pairs(
  emm_mass_time,
  adjust = "tukey"
)

emmeans(
  m_mass_time,
  pairwise ~ time | treatment * sex,
  adjust = "tukey"
)

```

## Reduced model: all 2-way interactions

```{r}
m_mass_2way <- lmer(
  body_mass ~
    time + treatment + sex +
    time:treatment +
    time:sex +
    treatment:sex +
    (1 | fish_id),
  data = mass_long,
  REML = FALSE
)

summary(m_mass_2way)

# Compare the models
anova(m_mass_2way, m_mass_time)
```

## Testing the 2-way interactions

```{r}
# 1. Remove time × treatment
m_mass_no_time_treatment <- update(
  m_mass_2way,
  . ~ . - time:treatment
)

anova(m_mass_no_time_treatment, m_mass_2way)


# 2. Remove time × sex
m_mass_no_time_sex <- update(
  m_mass_2way,
  . ~ . - time:sex
)

anova(m_mass_no_time_sex, m_mass_2way)


# 3. Remove treatment × sex
m_mass_no_treatment_sex <- update(
  m_mass_2way,
  . ~ . - treatment:sex
)

anova(m_mass_no_treatment_sex, m_mass_2way)
```

## Main effects model: the final model

```{r}
m_mass_additive <- lmer(
  body_mass ~ time + treatment + sex +
    (1 | fish_id),
  data = mass_long,
  REML = FALSE
)

summary(m_mass_additive)

# Diagnostics
res_mass <- simulateResiduals(
  fittedModel = m_mass_additive,
  n = 1000
)
plot(res_mass)
testDispersion(res_mass)
testOutliers(res_mass)
testUniformity(res_mass)

# Overall tests of fixed effects
anova(m_mass_additive, type = 3)

anova(m_mass_additive, m_mass_2way)
```

## Calculating the effect sizes (emmeans)

```{r}
# Estimated marginal means for time
emm_mass_time <- emmeans(m_mass_additive, ~ time)
emm_mass_time

# Before vs after LPS
pairs(emm_mass_time)


# Estimated marginal means for sex
emm_mass_sex <- emmeans(m_mass_additive, ~ sex)
emm_mass_sex

# Female vs male
pairs(emm_mass_sex)


# Estimated marginal means for treatment
emm_mass_treatment <- emmeans(m_mass_additive, ~ treatment)
emm_mass_treatment
```



## Plotting body-condition before and after injection across treatments

```{r}
# Prepare raw data for plotting
mass_long <- mass_long %>%
  mutate(
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    time = factor(
      time,
      levels = c("Before LPS", "After LPS")
    )
  )


# Estimated marginal means from final model
emm_mass_plot_df <- as.data.frame(
  emmeans(
    m_mass_additive,
    ~ time | treatment
  )
)

emm_mass_plot_df <- emm_mass_plot_df %>%
  mutate(
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    time = factor(
      time,
      levels = c("Before LPS", "After LPS")
    )
  )


# Colour palettes
pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)


# Plot
p_mass_lps <- ggplot() +

  # Raw paired individual trajectories
  geom_line(
    data = mass_long,
    aes(
      x = time,
      y = body_mass,
      group = fish_id
    ),
    alpha = 0.12,
    linewidth = 0.4,
    colour = "grey40"
  ) +

  # Raw observations
  geom_point(
    data = mass_long,
    aes(
      x = time,
      y = body_mass,
      colour = treatment_plot
    ),
    alpha = 0.35,
    size = 1.8,
    position = position_jitter(width = 0.05, height = 0)
  ) +

  # Estimated marginal means
  geom_line(
    data = emm_mass_plot_df,
    aes(
      x = time,
      y = emmean,
      group = treatment_plot,
      colour = treatment_plot
    ),
    linewidth = 1.4
  ) +

  # 95% CI
  geom_errorbar(
    data = emm_mass_plot_df,
    aes(
      x = time,
      ymin = lower.CL,
      ymax = upper.CL,
      colour = treatment_plot
    ),
    width = 0.08,
    linewidth = 1.1
  ) +

  geom_point(
    data = emm_mass_plot_df,
    aes(
      x = time,
      y = emmean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.1,
    colour = "black"
  ) +

  facet_wrap(~ treatment_plot) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    breaks = scales::pretty_breaks(6),
    expand = expansion(mult = c(0.02, 0.06))
  ) +

  labs(
    x = NULL,
    y = "Body mass (g)"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_mass_lps)

# Save

ggsave(
  filename = here("figures", "body_mass_before_after_LPS.pdf"),
  plot = p_mass_lps,
  width = 8,
  height = 4.8
)

ggsave(
  filename = here("figures", "body_mass_before_after_LPS.jpg"),
  plot = p_mass_lps,
  width = 8,
  height = 4.8,
  dpi = 600
)
```

## Body condition sorted by sex

```{r}
# Prepare raw data for plotting
mass_long <- mass_long %>%
  mutate(
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    ),
    time = factor(
      time,
      levels = c("Before LPS", "After LPS")
    )
  )


# Estimated marginal means from final model
emm_mass_sex_plot_df <- as.data.frame(
  emmeans(
    m_mass_additive,
    ~ time | sex
  )
)

emm_mass_sex_plot_df <- emm_mass_sex_plot_df %>%
  mutate(
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    ),
    time = factor(
      time,
      levels = c("Before LPS", "After LPS")
    )
  )


# Colour palettes
pastel_sex_cols <- c(
  "Female" = "#D8B4E2",
  "Male"   = "#A9CCE3"
)

dark_sex_cols <- c(
  "Female" = "#8E44AD",
  "Male"   = "#2874A6"
)


# Plot
p_mass_sex <- ggplot() +

  # Raw paired individual trajectories
  geom_line(
    data = mass_long,
    aes(
      x = time,
      y = body_mass,
      group = fish_id
    ),
    alpha = 0.12,
    linewidth = 0.4,
    colour = "grey40"
  ) +

  # Raw observations
  geom_point(
    data = mass_long,
    aes(
      x = time,
      y = body_mass,
      colour = sex_plot
    ),
    alpha = 0.35,
    size = 1.8,
    position = position_jitter(width = 0.05, height = 0)
  ) +

  # Model-estimated means
  geom_line(
    data = emm_mass_sex_plot_df,
    aes(
      x = time,
      y = emmean,
      group = sex_plot,
      colour = sex_plot
    ),
    linewidth = 1.4
  ) +

  # 95% confidence intervals
  geom_errorbar(
    data = emm_mass_sex_plot_df,
    aes(
      x = time,
      ymin = lower.CL,
      ymax = upper.CL,
      colour = sex_plot
    ),
    width = 0.08,
    linewidth = 1.1
  ) +

  # Estimated marginal means
  geom_point(
    data = emm_mass_sex_plot_df,
    aes(
      x = time,
      y = emmean,
      fill = sex_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.1,
    colour = "black"
  ) +

  # Separate panel for each sex
  facet_wrap(~ sex_plot) +

  scale_colour_manual(values = dark_sex_cols) +
  scale_fill_manual(values = pastel_sex_cols) +

  scale_y_continuous(
    breaks = scales::pretty_breaks(6),
    expand = expansion(mult = c(0.02, 0.06))
  ) +

  labs(
    x = NULL,
    y = "Body mass (g)"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_mass_sex)

# Save

ggsave(
  filename = here("figures", "body_mass_before_after_LPS_sex.pdf"),
  plot = p_mass_sex,
  width = 8,
  height = 4.8
)

ggsave(
  filename = here("figures", "body_mass_before_after_LPS_sex.jpg"),
  plot = p_mass_sex,
  width = 8,
  height = 4.8,
  dpi = 600
)

```

