---
title: "immunity"
author: "Gabriel Melhado"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Load libraries
pacman::p_load(car, dplyr, ggplot2, scales, here, tidyr, tidyverse, lme4,lmerTest,emmeans, performance, broom.mixed, DHARMa, ggpubr, psych, glmmTMB)
```

```{r}
# Read dataset
data <- read.csv("dataset.csv")
```


# Exploratory Data Analysis

```{r}
#Data structure
glimpse(data)
summary(data)
colSums(is.na(data))
```

```{r}
#Descriptive analysis
data %>%
  group_by(treatment) %>%
  reframe(across(where(is.numeric),
                 \(x) c(mean = mean(x, na.rm = TRUE),
                        sd = sd(x, na.rm = TRUE),
                        median = median(x, na.rm = TRUE),
                        min = min(x, na.rm = TRUE),
                        max = max(x, na.rm = TRUE))))
```


```{r}
# Checking data distribution

summary(data$total_wbc)
summary(data$n.l_ratio)


hist(data$total_wbc)
hist(data$n.l_ratio)
hist(data$refuge_duration_s)
hist(data$velocity_cm.s)

```

# Data Analysis: immune response

## Immunity * sex interaction
## Total_wbc: the total count of white blood cells circulating in the blood stream

```{r}
# Prepare variables
data$treatment <- factor(data$treatment,
                       levels = c("control", "low", "high"))

data$sex <- factor(data$sex)
data$tank_id <- factor(data$tank_id)
data$body_mass_sc <- scale(data$body_mass)

# Total WBC interaction model
m1_wbc_int <- glmmTMB(
  total_wbc ~ treatment * sex + body_mass_sc + (1 | tank_id),
  family = Gamma(link = "log"),
  data = data
)

summary(m1_wbc_int)

res_wbc_int <- simulateResiduals(m1_wbc_int)
plot(res_wbc_int)
testDispersion(res_wbc_int)
testOutliers(res_wbc_int)

Anova(m1_wbc_int, type = "III")

emmeans(
  m1_wbc_int,
  pairwise ~ treatment | sex,
  type = "response",
  adjust = "tukey"
)

emmeans(
  m1_wbc_int,
  pairwise ~ sex | treatment,
  type = "response",
  adjust = "tukey"
)
```

## Plotting: WBC x sex

```{r}
#---------------------------------------------------------
# Prepare labels
#---------------------------------------------------------

data <- data %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  )

#---------------------------------------------------------
# Colour palettes for sex
# Same as previous sex-split behavioural plots
#---------------------------------------------------------

sex_pastel_cols <- c(
  "Female" = "#D8B4E2",
  "Male"   = "#A9CCE3"
)


sex_dark_cols <- c(
  "Female" = "#8E44AD",
  "Male"   = "#2874A6"
)

pd <- position_dodge(width = 0.75)

pj <- position_jitterdodge(
  jitter.width = 0.12,
  dodge.width = 0.75
)

#=========================================================
# 1) TOTAL WBC
#=========================================================

#---------------------------------------------------------
# Estimated marginal means
#---------------------------------------------------------

emm_wbc_int_df <- as.data.frame(
  emmeans(
    m1_wbc_int,
    ~ sex | treatment,
    type = "response"
  )
)

cn <- names(emm_wbc_int_df)

mean_col <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_wbc_int_df <- emm_wbc_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

#---------------------------------------------------------
# Plot total WBC
#---------------------------------------------------------

p_wbc_sex <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = total_wbc,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = total_wbc,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_wbc_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_wbc_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    breaks = scales::pretty_breaks(8),
    expand = expansion(mult = c(0, 0.05))
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Total white blood cells",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_wbc_sex)

ggsave(
  filename = here("figures", "total_wbc_treatment_sex_side_by_side.pdf"),
  plot = p_wbc_sex,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "total_wbc_treatment_sex_side_by_side.jpg"),
  plot = p_wbc_sex,
  width = 7,
  height = 5,
  dpi = 600
)
```

## Main effects model

```{r}
m1_wbc <- glmmTMB(
  total_wbc ~ treatment + sex + body_mass_sc + (1 | tank_id),
  family = Gamma(link = "log"),
  data = data
)

summary(m1_wbc)

simulationOutput <- simulateResiduals(m1_wbc)
plot(simulationOutput)
testDispersion(simulationOutput)
testOutliers(simulationOutput)

Anova(m1_wbc, type = "II")
emm_wbc <- emmeans(
    m1_wbc,
    ~ treatment,
    type = "response"
)

anova(m1_wbc, m1_wbc_int)

emm_wbc

pairs(emm_wbc, adjust = "tukey")

```

## Plotting
## total_wbc

```{r}
# Prepare treatment labels for plotting
data <- data %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  )


# Estimated marginal means (response scale)
emm_wbc_df <- as.data.frame(
  emmeans(
    m1_wbc,
    ~ treatment,
    type = "response"
  )
)

print(emm_wbc_df)


# Standardise column names
cn <- names(emm_wbc_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_wbc_df <- emm_wbc_df %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )


# Colour palette
pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)


# Plot
p_wbc <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = total_wbc,
      fill = treatment_plot
    ),
    trim = FALSE,
    alpha = 0.25,
    colour = NA
  ) +

  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = total_wbc,
      colour = treatment_plot
    ),
    width = 0.18,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_wbc_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = treatment_plot
    ),
    width = 0.25,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_wbc_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    limits = c(0, 80),
    breaks = seq(0, 80, by = 10),
    expand = expansion(mult = c(0, 0.02))
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Total white blood cells"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    plot.title = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_wbc)

# Save
ggsave(
  here("figures", "total_wbc.pdf"),
  p_wbc,
  width = 6,
  height = 5
)

ggsave(
  here("figures", "total_wbc.jpg"),
  p_wbc,
  width = 6,
  height = 5,
  dpi = 600
)
```

## Immunity * sex interaction
## N.L_ratio: neutrophil-to-lymphocye ratio. Two of the most abundant white blood cells.

```{r}
# N:L ratio interaction model
m1_nl_int <- glmmTMB(
  n.l_ratio ~ treatment * sex + body_mass_sc + (1 | tank_id),
  family = Gamma(link = "log"),
  data = data
)

summary(m1_nl_int)

res_nl_int <- simulateResiduals(m1_nl_int)
plot(res_nl_int)
testDispersion(res_nl_int)
testOutliers(res_nl_int)

Anova(m1_nl_int, type = "III")

emmeans(
  m1_nl_int,
  pairwise ~ treatment | sex,
  type = "response",
  adjust = "tukey"
)

emmeans(
  m1_nl_int,
  pairwise ~ sex | treatment,
  type = "response",
  adjust = "tukey"
)
```

## Plotting N.L ratio x Sex

```{r}
# Estimated marginal means
emm_nl_int_df <- as.data.frame(
  emmeans(
    m1_nl_int,
    ~ sex | treatment,
    type = "response"
  )
)

cn <- names(emm_nl_int_df)

mean_col <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_nl_int_df <- emm_nl_int_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    sex = factor(sex, levels = c("female", "male")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    ),
    sex_plot = factor(
      sex,
      levels = c("female", "male"),
      labels = c("Female", "Male")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )


# Plot N:L ratio
p_nl_sex <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = n.l_ratio,
      fill = sex_plot,
      group = interaction(treatment_plot, sex_plot)
    ),
    position = pd,
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_point(
    data = data,
    aes(
      x = treatment_plot,
      y = n.l_ratio,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pj,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_nl_int_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = sex_plot,
      group = sex_plot
    ),
    position = pd,
    width = 0.18,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_nl_int_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = sex_plot,
      group = sex_plot
    ),
    position = pd,
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = sex_dark_cols) +
  scale_fill_manual(values = sex_pastel_cols) +

  scale_y_continuous(
    breaks = scales::pretty_breaks(8),
    expand = expansion(mult = c(0, 0.05))
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Neutrophil:lymphocyte ratio",
    fill = "Sex",
    colour = "Sex"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "top",
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 12),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_nl_sex)

ggsave(
  filename = here("figures", "nl_ratio_treatment_sex_side_by_side.pdf"),
  plot = p_nl_sex,
  width = 7,
  height = 5
)

ggsave(
  filename = here("figures", "nl_ratio_treatment_sex_side_by_side.jpg"),
  plot = p_nl_sex,
  width = 7,
  height = 5,
  dpi = 600
)

```


## Main effects model

```{r}
# 1) Prepare variables
data <- data %>%
  mutate(
    treatment = factor(
      treatment,
      levels = c("control", "low", "high")
    ),
    sex = factor(sex),
    tank_id = factor(tank_id),
    body_mass_sc = as.numeric(scale(body_mass)),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  )


# 2) Model: Gamma GLMM

m1_nl <- glmmTMB(
  n.l_ratio ~ treatment + sex + body_mass_sc + (1 | tank_id),
  family = Gamma(link = "log"),
  data = data
)

summary(m1_nl)


# 3) Diagnostics

res_nl <- simulateResiduals(m1_nl)
plot(res_nl)
testDispersion(res_nl)
testOutliers(res_nl)

# 4) Type II ANOVA
Anova(m1_nl, type = "II")

anova(m1_nl, m1_nl_int)


# 5) Estimated marginal means
emm_nl <- emmeans(
  m1_nl,
  ~ treatment,
  type = "response"
)

emm_nl

pairs(emm_nl, adjust = "tukey")
```

## Plottiing: n.l_ratio

```{r}
# Estimated marginal means dataframe

emm_nl_df <- as.data.frame(emm_nl)

print(emm_nl_df)

# Standardise column names
cn <- names(emm_nl_df)

mean_col  <- dplyr::first(cn[cn %in% c("response", "rate", "emmean")])

lower_col <- dplyr::first(
  cn[grepl("^lower\\.CL$|^asymp\\.LCL$|^LCL$|^lower$", cn)]
)

upper_col <- dplyr::first(
  cn[grepl("^upper\\.CL$|^asymp\\.UCL$|^UCL$|^upper$", cn)]
)

emm_nl_df <- emm_nl_df %>%
  mutate(
    treatment = factor(treatment, levels = c("control", "low", "high")),
    treatment_plot = factor(
      treatment,
      levels = c("control", "low", "high"),
      labels = c("Control", "Low", "High")
    )
  ) %>%
  rename(
    mean  = !!mean_col,
    lower = !!lower_col,
    upper = !!upper_col
  )

# Colour palette

pastel_cols <- c(
  "Control" = "#A8D5BA",
  "Low"     = "#F9E79F",
  "High"    = "#F5B7B1"
)

dark_cols <- c(
  "Control" = "#2E8B57",
  "Low"     = "#C9A227",
  "High"    = "#B22222"
)

# Plot


p_nl <- ggplot() +

  geom_violin(
    data = data,
    aes(
      x = treatment_plot,
      y = n.l_ratio,
      fill = treatment_plot
    ),
    trim = FALSE,
    alpha = 0.25,
    colour = NA,
    linewidth = 0
  ) +

  geom_jitter(
    data = data,
    aes(
      x = treatment_plot,
      y = n.l_ratio,
      colour = treatment_plot
    ),
    width = 0.18,
    alpha = 0.55,
    size = 2
  ) +

  geom_errorbar(
    data = emm_nl_df,
    aes(
      x = treatment_plot,
      ymin = lower,
      ymax = upper,
      colour = treatment_plot
    ),
    width = 0.25,
    linewidth = 1.3
  ) +

  geom_point(
    data = emm_nl_df,
    aes(
      x = treatment_plot,
      y = mean,
      fill = treatment_plot
    ),
    shape = 21,
    size = 5,
    stroke = 1.2,
    colour = "black"
  ) +

  scale_colour_manual(values = dark_cols) +
  scale_fill_manual(values = pastel_cols) +

  scale_y_continuous(
    breaks = scales::pretty_breaks(8),
    expand = expansion(mult = c(0, 0.02))
  ) +

  labs(
    x = "Fluoxetine treatment",
    y = "Neutrophil:lymphocyte ratio"
  ) +

  theme_bw(base_size = 14) +

  theme(
    panel.grid = element_blank(),
    legend.position = "none",
    plot.title = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13)
  )

print(p_nl)


# Save
ggsave(
  filename = here("figures", "nl_ratio.pdf"),
  plot = p_nl,
  width = 6,
  height = 5
)

ggsave(
  filename = here("figures", "nl_ratio.jpg"),
  plot = p_nl,
  width = 6,
  height = 5,
  dpi = 600
)
```

